The Forward March of Categorical Tolerance in the United States\thanks{Reserved for acknowledgments}

### 2025-11-15 manuscript.tex

\documentclass[12pt]{article}
\usepackage[left=1.15in, right=1.15in, top=1in, bottom=1.5in]{geometry}
\usepackage{natbib}
\usepackage{verbatim}
\usepackage{setspace}
\usepackage{booktabs}
\usepackage{caption}
\usepackage{graphicx}
\usepackage{subcaption}
\bibpunct[: ]{(}{)}{;}{a}{}{,} 
\usepackage{authblk}
\renewcommand\Authfont{\normalsize}
\renewcommand\Affilfont{\footnotesize}

@@

\newpage
\begin{abstract}
This paper updates the empirical picture of Categorical Tolerance (CT), namely, the pattern of refusing to report dislikes across cultural genres, for the third decade of the twenty-first century in the United States. Analyzing recent survey data from two platforms, I find that CT has continued its march among Americans, reaching approximately 20\% of respondents. The analysis confirms earlier-observed demographic trends, showing that CT is strongly associated with younger cohorts and non-white individuals. However, I also find that individuals reporting the highest educational attainment are now overrepresented among categorical tolerants, suggesting that CT may increasingly function as an elite cultural strategy consistent with contemporary forms of status display, signaling openness and refusal to refuse. Furthermore, the study uses political ideology as a predictor of symbolic exclusion, revealing that while the odds of being a CT itself are not strongly polarized, the inclination toward symbolic exclusion among non-CTs is, with conservatives significantly more likely to express a greater volume of cultural dislikes than liberals, showing that political ideology, alongside age and ethnoracial identity, is a central determinant of cultural exclusion in the contemporary context. 
\end{abstract}
\newpage
\section*{Introduction}
Ever since Richard Peterson introduced the felicitous term ``patterns of cultural choice'' into sociological discourse more than four decades ago \citep{peterson1983patterns-f85}, sociologists have been busy empirically uncovering and debating the sociological reality, drivers, meaning, and correlates of various such patterns. These include, most famously and influentially, the Petersonian cultural \textit{omnivore} \citep{peterson1996changing-967, goldberg2011mapping-77a, warde2008omnivorous-e77} along with their evil twin the culturally exclusive \textit{univore} \citep{bryson1997what-933, peterson1992understanding-ce3, tonkovi2025globallocal-96f}, but also culturally \textit{voracious} consumers \citep{sullivan2007omnivore-018}, \textit{poly-purists} and \textit{poly-mixers} \citep{goldberg2016what-4e3, poecke2018pure-873}, \textit{decisives} and \textit{indecisives} \citep{sonnett2004musical-961}, and even plain old \textit{snobs} engaging in (presumably) Bourdieusian boundary-drawing distinction \citep{atkinson2011context-d44, nault2021social-1bf}, among many others.

>>>>

\documentclass[12pt]{article}
\usepackage[left=1.15in, right=1.15in, top=1in, bottom=1.5in]{geometry}
\usepackage{natbib}
\usepackage{verbatim}
\usepackage{setspace}
\usepackage{booktabs}
\usepackage{caption}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{url}
\bibpunct[: ]{(}{)}{;}{a}{}{,} 
\usepackage{authblk}
\renewcommand\Authfont{\normalsize}
\renewcommand\Affilfont{\footnotesize}

@@

\newpage
\begin{abstract}
This paper updates the empirical picture of Categorical Tolerance (CT), namely, the pattern of refusing to report dislikes across cultural genres, for the third decade of the twenty-first century in the United States. Analyzing recent survey data from two platforms, I find that CT has continued its march among Americans, reaching approximately 20\% of respondents. The analysis confirms earlier-observed demographic trends, showing that CT is strongly associated with younger cohorts and non-white individuals. However, I also find that individuals reporting the highest educational attainment are now overrepresented among categorical tolerants, suggesting that CT may increasingly function as an elite cultural strategy consistent with contemporary forms of status display, signaling openness and refusal to refuse. Furthermore, the study uses political ideology as a predictor of symbolic exclusion, revealing that while the odds of being a CT itself are not strongly polarized, the inclination toward symbolic exclusion among non-CTs is, with conservatives significantly more likely to express a greater volume of cultural dislikes than liberals, showing that political ideology, alongside age and ethnoracial identity, is a central determinant of cultural exclusion in the contemporary context.

\textbf{Reproducibility Package:} Data files and \textit{R} code (in Quarto Markdown format) necessary to reproduce all of the analyses, tables, and figures reported in the paper can be found at the following GitHub repo: \url{https://github.com/olizardo/sociological-science-categorical-tolerance-followup}.
\end{abstract}
\newpage
\section*{Introduction}
Ever since Richard Peterson introduced the felicitous term ``patterns of cultural choice'' into sociological discourse more than four decades ago \citep{peterson1983patterns-f85}, sociologists have been busy empirically uncovering and debating the sociological reality, drivers, meaning, and correlates of various such patterns. These include, most famously and influentially, the Petersonian cultural \textit{omnivore} \citep{peterson1996changing-967, goldberg2011mapping-77a, warde2008omnivorous-e77} along with their evil twin the culturally exclusive \textit{univore} \citep{bryson1997what-933, peterson1992understanding-ce3, tonkovi2025globallocal-96f}, but also culturally \textit{voracious} consumers \citep{sullivan2007omnivore-018}, \textit{poly-purists} and \textit{poly-mixers} \citep{goldberg2016what-4e3, poecke2018pure-873}, \textit{decisives} and \textit{indecisives} \citep{sonnett2004musical-961}, and even plain old \textit{snobs} engaging in (presumably) Bourdieusian boundary-drawing distinction \citep{atkinson2011context-d44, nault2021social-1bf}, among many others.
